---
title: "simpleRegress"
author: "Isabela Bercian"
date: "17 April 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

According to the graph, the thickness of a bears neck is correlated to its overall weight. The thicker a bear's neck, the more it weighs. We can see this by observing the upward trend of the regression line. 

## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
mHEADLEN <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(mHEADLEN)
summary.lm(mHEADLEN)
```

The residual standard error is 43.91 and our adjusted R-squared is 0.69. With this much data, the data models in order from best to first are WEIGHT explained by NECK, followed by WEIGHT explained by HEADLEN.

```{r}
mHEADWTH <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(mHEADLEN)
summary.lm(mHEADWTH)
```

Now, in order from best to worst according to RSE: 
1. WEIGHT explained by NECK at 43.91  
2. WEIGHT explained by HEADLEN at 67.81  
3. WEIGHT explained by HEADWTH at 76.41  

```{r}
mLENGTH <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(mLENGTH)
summary.lm(mLENGTH)
```


Now, in order from best to worst according to RSE:  
1. WEIGHT explained by NECK at 43.91  
2. WEIGTH explained by LENGTH at 61.83  
3. WEIGHT explained by HEADLEN at 67.81  
4. WEIGHT explained by HEADWTH at 76.41  
 